---
title: "Data Preprocessing"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
births <- read.csv("chds_births.csv")
meth.names <- c('Caucasian','Caucasian','Caucasian','Caucasian','Caucasian','Caucasian', 'Mexican', 'African-American', 'Asian', 'Mixed', 'Other')
med.names <- c('elementary', 'middle', 'hs', 'hs + trade', 'hs + college', 'college', 'trade', 'unclear')
feth.names <- c('Caucasian','Caucasian','Caucasian','Caucasian','Caucasian','Caucasian', 'Mexican', 'African-American', 'Asian', 'Mixed', 'Other')
fed.names <- c('elementary', 'middle', 'hs', 'hs + trade', 'hs + college', 'college', 'trade', 'unclear')
marital.names <- c(NA, 'married', 'separated', 'divorced', 'widowed', 'never married')
income.names <- c('<2500', '2500-4999', '5000-7499', '7500-9999', '10000-12499', '12500-14999', '15000-17499', '20000-22499', '>22500')
smoke.names <- c('never', 'now', 'until pregnancy', 'used to')
time.names <- c('never', 'still smokes', 'during pregnancy', 'less than a year', '1-2yrs', '2-3yrs', '3-4yrs', '5-9yrs', '10+yrs', 'quit - unknown then')
number.names <- c('never', '1-4', '5-9', '10-14', '15-19', '20-29', '30-39', '40-60', '>60', 'smoked, amount unknown')

births$meth <- meth.names[births$meth + 1]
births$feth<- feth.names[births$feth + 1]
births$fed <- fed.names[births$fed + 1]
births$marital <- marital.names[births$marital+1]
births$income <- income.names[births$income + 1]
births$smoke <- smoke.names[births$smoke + 1]
births$time <- time.names[births$time + 1]
births$number <- number.names[births$number + 1]
summary(births)
```
## Initial Model

### Variables to definitely include
\begin{enumerate}
\item gestation
\item parity
\item time
\item number
\item smoke
\item martial
\item fed
\item med
\end{enumerate}

### Variables to consider including
\begin{enumerate}
\item meth
\item feth
\item mage/fage (not both -- correlated)
\item mht/mwt (slightly correlated so probably not both -- mwt might be better)
\end{enumerate}

### Non-linear effects/ other modifications to covariates
\begin{itemize}
\item change grouping of smoke: group 0 and 3 together; 1 and 2 together to form "never/used to" and "now/until pregnancy"
\item change grouping of med/fed: (0, 1, 7) becomes group "no highschool/ highschool unclear", (3, 6) -> trade, (4, 5) -> college [the latter 2 groupings are relevant for fed more than med]
\item income* (have to fix with imputation first)
\end{itemize}


### Forward Selection using just the variables in "to include"
```{r}
keeps <- c("wt", "gestation", "parity", "time", "number", "smoke", "marital", "fed", "med", "mwt", "mht")
birth.data <- births[keeps]
birth.data <- na.omit(birth.data)
print(dim(birth.data))
M0 <- lm(wt ~ 1, data = birth.data)
Mmax <- lm(wt ~ gestation + parity + time + number+ smoke + fed, data=birth.data)
Mstart <- lm(wt ~ gestation + parity + smoke + fed, data=birth.data)
MgestPar <- lm(wt ~ gestation + parity + I(gestation*parity) + smoke + fed, data = birth.data)
Mbio1 <- lm(wt ~ gestation + parity + mwt + mht + I(mwt*mht^2) + I(gestation*parity) + time, data = birth.data)
Mbio2 <- lm(wt ~ gestation + parity + mwt + mht + I(mwt*mht^2) + I(gestation*parity) + smoke, data = birth.data)
Mbio3 <- lm(wt ~ gestation + parity + mwt + mht + I(mwt*mht^2) + I(gestation*parity) + number, data = birth.data)
MbioBMI <- lm(wt ~ gestation + parity + mwt + mht + I(mwt*703/mht^2) + I(gestation*parity) + number, data = birth.data)
bmi <- birth.data$mwt*703/(birth.data$mht)^2
bmiLB <- bmi - 18.5
bmiUB <- 24.9 - bmi
birth.data$bmiLB <- bmiLB
birth.data$bmiUB <- bmiUB
MbioBMICenteredUB <- lm(wt ~ gestation + parity + mwt + mht + bmiUB + number, data = birth.data)
MbioBMICenteredLB <- lm(wt ~ gestation + parity + mwt + mht + bmiLB + number, data = birth.data)
MbioBMICentered <- lm(wt ~ gestation + parity + I(gestation*parity) + bmiUB + number, data = birth.data)
ntot <- dim(birth.data)[1]
ntrain <- 1000
set.seed(5)
train.ind <- sample(ntot, ntrain)

M0 <- update(M0, subset = train.ind)
Mmax <- update(Mmax, subset = train.ind)
Mstart <- update(Mstart, subset = train.ind)
```
```{r}
# forward selection
Mfwd <- step(object = M0, # starting point model
scope = list(lower = M0, upper = Mmax), # smallest and largest model
direction = "forward", 
trace = FALSE) # trace prints out information
print(Mfwd$call)
```
```{r}
# backward elimiation
Mback <- step(object = Mmax, # starting point model
scope = list(lower = M0, upper = Mmax),
direction = "backward", trace = FALSE)
print(Mback$call)
```
```{r}
# stepwise selection (both directions)
Mstep <- step(object = Mstart,
scope = list(lower = M0, upper = Mmax),
direction = "both", trace = FALSE)
print(Mstep$call)
```

The MSPE for training set of 1000, seed=5, omit NA is 
```{r}
print(Mfwd$call)
print(sum((birth.data$wt[-train.ind] - predict(Mfwd, newdata = birth.data[-train.ind,]))^2))
print(Mback$call)
print(sum((birth.data$wt[-train.ind] - predict(Mback, newdata = birth.data[-train.ind,]))^2))
print(Mstep$call)
print(sum((birth.data$wt[-train.ind] - predict(Mstep, newdata = birth.data[-train.ind,]))^2))
print(Mstart$call)
print(sum((birth.data$wt[-train.ind] - predict(Mstart, newdata = birth.data[-train.ind,]))^2))
print(MgestPar$call)
print(sum((birth.data$wt[-train.ind] - predict(MgestPar, newdata = birth.data[-train.ind,]))^2))
print(Mbio1$call)
print(sum((birth.data$wt[-train.ind] - predict(Mbio1, newdata = birth.data[-train.ind,]))^2))

print(Mbio2$call)
print(sum((birth.data$wt[-train.ind] - predict(Mbio2, newdata = birth.data[-train.ind,]))^2))

print(Mbio3$call)
print(sum((birth.data$wt[-train.ind] - predict(Mbio3, newdata = birth.data[-train.ind,]))^2))

print(MbioBMI$call)
print(sum((birth.data$wt[-train.ind] - predict(MbioBMI, newdata = birth.data[-train.ind,]))^2))

print(MbioBMICentered$call)
print(sum((birth.data$wt[-train.ind] - predict(MbioBMICentered, newdata = birth.data[-train.ind,]))^2))

print(MbioBMICenteredLB$call)
print(sum((birth.data$wt[-train.ind] - predict(MbioBMICenteredLB, newdata = birth.data[-train.ind,]))^2))

print(MbioBMICenteredUB$call)
print(sum((birth.data$wt[-train.ind] - predict(MbioBMICenteredUB, newdata = birth.data[-train.ind,]))^2))
```
