---
title: "Data Imputation for Income"
output: pdf_document
---

```{r  include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=3)
knitr::opts_chunk$set(echo=FALSE)
options(warn = -1)
```
The functions in this file are dependent on the libraries VIM and mice
```{r include=FALSE}
#Reading in the data 
fdata <- read.csv("chds_births.csv")
#Calculat the number of missing valuesin each column 
na_count <- sapply(fdata, function(y) sum(length(which(is.na(y)))))
na.count <- data.frame(na_count)
#Check the columns that have more than 10% of the data missing 
count <- sapply(fdata, function(y) length(y))
na_percent <- (na_count/count)*100
na_percent <- data.frame(na_percent)
```
Visualizing the missing data
```{r  include = FALSE}
library(VIM)
```
```{r results = 'hide' }
miss_plot <- aggr(fdata, col=c('navyblue','yellow'),
                    numbers=TRUE, sortVars=TRUE,
                    labels=names(fdata), cex.axis=.7,
                    gap=3, ylab=c("Missing data","Pattern"))
```

Looking the the above data frame, we might want to remove or impute values into the following features of the data, fwt : Fathers Weight,fht : Fathers Height, income 
Since imputting may not work for fathers height and weight, we will remove those features from our data. Logically, it makes sense for the baby's weight to not depend on these features.
```{r include=FALSE}
fdata$fht <- NULL
fdata$fwt <- NULL 
#colnames(fdata)
```
Using the mice library, we perform imputation for the income covariate
```{r include=FALSE}
library(mice)
#Imputting the data with "pmm" method as referenced from "https://stefvanbuuren.name/mice/"
imp <- mice(fdata, m = 5, meth = 'pmm', seed = 500)
```

```{r echo = FALSE}
densityplot(imp)
```

Blue - observed values   Magenta - imputed values 

From the density plot, we can say that imputed values are indeed plausible for the income covariate.

All the imputed data sets 
```{r}
imp_1 <- data.frame(complete(imp,1))
imp_2 <- data.frame(complete(imp,2))
imp_3 <- data.frame(complete(imp,3))
imp_4 <- data.frame(complete(imp,4))
imp_5 <- data.frame(complete(imp,5))
```
Since the imputted values for income for all the imputed data sets follow a similar distribution, we choose imputed dataset 1, since it
has the smallest difference in the means when compared the observed data
```{r}
cm1 <- sapply(fdata, mean, na.rm = T) - sapply(imp_1, mean)
cm2 <- sapply(fdata, mean, na.rm = T) - sapply(imp_2, mean)
cm3 <- sapply(fdata, mean, na.rm = T) - sapply(imp_3, mean)
cm4 <- sapply(fdata, mean, na.rm = T) - sapply(imp_4, mean)
cm5 <- sapply(fdata, mean, na.rm = T) - sapply(imp_5, mean)
sum_imp_data <- c(sum(cm1), sum(cm2), sum(cm3), sum(cm4), sum(cm5))
abs(sum_imp_data)
```
We choose to include the imputed income data from imputed dataset 1 into our original dataset and then carry our model diagnostics and selection from there. 
```{r}
#Get sample 1 income values
imputed_income <- imp_1$income
fdata$income <- imputed_income
```


