---
title: "Variable Exploration"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Covariates
\begin{enumerate}
\item \textbf{wt}
\begin{itemize}
\item The birth weight (ounces).
\item continuous
\end{itemize}
\item \textbf{gestation}
\begin{itemize}
\item The length of the gestation period (days).
\item continuous
\end{itemize}
\item\textbf{parity}
\begin{itemize}
\item The total number of previous pregnancies (including fetal deaths and still births).
\item discrete, not categorical
\end{itemize}
\item\textbf{meth}
\begin{itemize}
\item The self-reported ethnicity of the mother: 
\begin{itemize}
\item 0-5 = Caucasian, 
\item 6 = Mexican, 
\item 7 = African-American,
\item 8 = Asian, 
\item 9 = Mixed, 
\item 10= Other
\end{itemize}
\item categorical
\end{itemize}
\item\textbf{mage}
\begin{itemize}
\item The mother’s age at termination of pregnancy (years).
\item theoretically continuous, in practice probably discrete *double check this*
\end{itemize}
\item\textbf{med}
\begin{itemize}
\item The mother’s education: 
\begin{itemize}
\item 0 = elementary school
\item 1 = middle school
\item 2 = high school
\item 3 = high school + trade school, 
\item 4 = high school + some college, 
\item 5 = college graduate,
\item 6 = trade school, 
\item 7 = highschool unclear.
\end{itemize}
\item categorical
\end{itemize}
\item\textbf{mht}
\begin{itemize}
\item The mother’s height (inches).
\item continuous
\end{itemize}
\item\textbf{mwt}
\begin{itemize}
\item The mother’s pregnancy weight (pounds).
\item continuous
\end{itemize}
\item\textbf{feth}
\begin{itemize}
\item The father’s ethnicity (same coding as meth ).
\begin{itemize}
\item 0-5 = Caucasian, 
\item 6 = Mexican, 
\item 7 = African-American,
\item 8 = Asian, 
\item 9 = Mixed, 
\item 10= Other
\end{itemize}
\item categorical
\end{itemize}
\item\textbf{fage}
\begin{itemize}
\item The father’s age at end of pregnancy (years).
\item theoretically continuous, probably discrete in practice *double check*
\end{itemize}
\item\textbf{fed}
\begin{itemize}
\item The father’s education (same coding as med ).
\begin{itemize}
\item 0 = elementary school
\item 1 = middle school
\item 2 = high school
\item 3 = high school + trade school, 
\item 4 = high school + some college, 
\item 5 = college graduate,
\item 6 = trade school, 
\item 7 = highschool unclear.
\end{itemize}
\item categorical
\end{itemize}
\item\textbf{fht}
\begin{itemize}
\item The father’s height (inches).
\item continuous
\end{itemize}
\item\textbf{fwt}
\begin{itemize}
\item the father’s weight (pounds).
\item continuous
\end{itemize}
\item\textbf{marital}
\begin{itemize}
\item The mother’s marital status: 
\begin{enumerate}
\item married, 
\item legally separated
\item divorced
\item widowed,
\item never married.
\end{enumerate}
\item categorical
\end{itemize}
\item\textbf{income}
\begin{itemize}
\item The family yearly income in \$2500 (USD) increments:
\begin{itemize}
\item 0 = under 2500
\item 1 = 2500-4999
\item 2 = 5000-7499
\item 3 = 7500-9999
\item 4 = 10000-12499
\item 5 = 12500-14999
\item 6 = 15000-17499
\item 7 = 17500-19999
\item 8 = 20000-22499
\item 9 = over 22500.
\end{itemize}
\item categorical
\end{itemize}
\item\textbf{smoke}
\begin{itemize}
\item Does the mother smoke at time of pregnancy?
\begin{itemize}
\item 0 = never
\item 1 = smokes now
\item 2 = until pregnancy
\item 3 = used to, not anymore.
\end{itemize}
\item categorical
\end{itemize}
\item\textbf{time}
\begin{itemize}
\item Time since the mother quit smoking before pregnancy:
\begin{itemize}
\item 0 = never smoked, 
\item 1 = still smokes, 
\item 2 = during pregnancy
\item 3 = less than a year,
\item 4 = 1-2 years, 
\item 5 = 2-3 years, 
\item 6 = 3-4 years, 
\item 7 = 5-9 years, 
\item 8 = more than 10 years,
\item 9 = quit but don’t know when.
\end{itemize}
\item categorical
\end{itemize}
\item\textbf{number}
\begin{itemize}
\item Number of cigarettes smoked per day by mother when she was smoking:
\begin{itemize}
\item 0 = never smoked,
\item 1 = 1-4, 
\item 2 = 5-9, 
\item 3 = 10-14,
\item 4 = 15-19,
\item 5 = 20-29, 
\item 6 = 30=39,
\item 7 = 40-60, 
\item 8 = more than 60, 
\item 9 = smoked but don’t know how much.
\end{itemize}
\item categorical
\end{itemize}
\end{enumerate}
```{r}
cat.inds <- c(4,6,9,11,14,15,16,17,18)
discr.inds <- c(3, 5, 10 )
cts.inds <- c(1, 2,7,8,12,13)
births <- read.csv("chds_births.csv")
```
```{r}
pairs(births[, c(1, 4, 6, 9, 11, 14)], pch=19)
```

Variables to include: marital, fed, med, maybe meth, maybe feth
```{r}
pairs(births[, c(1, 15, 16, 17, 18)], pch=19)
```

time, number, income, maybe smoke (should try grouping 0 and 3 together; 1 and 2 together)

```{r}
pairs(births[, c(1, 3, 5, 10)], pch=19)
```

Variables to include: parity, maybe mage or fage (note that mage and fage are correlated)

```{r}
pairs(births[, c(1, 2,7,8)], pch=19)
```

Variables to include: gestation, maybe mht/mwt (note that these are slightly correlated)

```{r}
pairs(births[, c(1, 12, 13)], pch=19)
```

fht and fwt do not seem to be good variables to include (and are definitely correlated)

# Missing Data
Note that 5-10\% of observations = ```r (dim(births))[1]*0.05```-```r (dim(births))[1]*0.10```

```{r}
summary(births)
```
Variables to remove due to missing data:
\begin{itemize}
\item fht (492)
\item fwt (499)
\item income (124)
\end{itemize}

# Ideas for Initial Model

## Variables to include
\begin{enumerate}
\item gestation
\item parity
\item time
\item number
\item smoke
\item martial
\item fed
\item med
\end{enumerate}

## Variables to consider including
\begin{enumerate}
\item meth
\item feth
\item mage/fage (not both -- correlated)
\item mht/mwt (slightly correlated so probably not both -- mwt might be better)
\end{enumerate}

## Non-linear effects/ other modifications to covariates
\begin{itemize}
\item change grouping of smoke: group 0 and 3 together; 1 and 2 together to form "never/used to" and "now/until pregnancy"
\item change grouping of med/fed: (0, 1, 7) becomes group "no highschool/ highschool unclear", (3, 6) -> trade, (4, 5) -> college [the latter 2 groupings are relevant for fed more than med]
\item income* (have to fix with imputation first)
\end{itemize}
